---
title: "Computer exercise - Market and Systemic Risk Management"
author: "A. Schmidt"
date: "June, 2021"
output: 
  pdf_document:
    extra_dependencies: ["float"]
header-includes:
- \usepackage{mathtools}
urlcolor: blue
linkcolor: red
---


# Introduction

This is the computational appendix with the analysis made for the Financial Report at the Pokè Bank Johto Region regarding the investment in the Rocket S.A. portfolio.

# Dependency installation

```{r, message = FALSE, warning = FALSE, echo = TRUE}
rm(list=ls())        # Clean memory

# Verify if a package is installed, if not, download and install before loading. 
chooseCRANmirror(graphics = FALSE, ind = 10)
if (!require("pacman")) install.packages("pacman")
pacman::p_load(reshape2, ggplot2, latex2exp, gridExtra, readxl, 
               RColorBrewer, grid, knitr, zoo, scales, evir)

# Shapper is a wrapper for the Python package with the Shap values. 
# In case you don't have it on Python, you need to install running the following line
#shapper::install_shap()
```

```{r, message = FALSE, warning = FALSE, echo = TRUE}
# Prevents code from getting out of the page
## Works with almost everything except urls and strings.
## Last options hold the position of figures
opts_chunk$set(tidy.opts=list(width.cutoff=50),tidy=TRUE, fig.pos = "!H", out.extra = "")

# Set a seed
set.seed(6969)

# Choosing a color palette
cores <- brewer.pal(8, "Dark2")

# Set a working directory
setwd("C:\\Users\\aisha\\OneDrive\\Documentos\\Mestrado Tinbergen\\Year 2\\Block 05\\Market and Systemic Risk\\Computational Assignment\\")
```


# Data preparation


```{r, eval = TRUE, message = FALSE, warning = FALSE}
url      <- "https://raw.githubusercontent.com/aishameriane/SystemicRisk/main/data.csv"
dfData01 <- read.csv2(url, sep = ",", dec = ".", header = TRUE)
dfData01 <- dfData01[, c("Market", "Schmidt")]

ending   <- as.Date("2021-06-25")
# Had to cut one day to make vectors with the same size
starting <- as.Date(as.Date("2021-06-25")-2516)
Date     <- as.Date(starting:ending, origin=starting)


Date <- seq(starting, ending, by = "1 days")
dfData01 <- cbind(Date, dfData01)
```


# Exploring data for heavy tails

```{r, eval = TRUE, message = FALSE, warning = FALSE}
# Data from recessions available at: https://eabcn.org/dc/chronology-euro-area-business-cycles
recessions_1 <- c("2019Q4")

recessions_2 <- c("2021Q3")

recessions.trim <- data.frame(1, 1,1, 
                              as.Date(as.yearqtr(recessions_1)), 
                              as.Date(as.yearqtr(recessions_2)))
names(recessions.trim) <- c("DATE", "Variable", "Value", "Peak", "Trough")


dfData02 <- melt(dfData01, id.vars = "Date")
names(dfData02) <- c("DATE", "Variable", "Value")

p0 <- ggplot(dfData02[which(dfData02$Variable == "Schmidt"), ]) +
        geom_line(aes(x = DATE, y = Value, color = Variable), alpha = 0.9)+
        labs(y = "Return", x= "", color = "Series") +
        scale_x_date(date_breaks = "12 months", labels=date_format("%Y"))+ 
        scale_colour_manual(values = c("#A43E53")) +
        theme_bw() +
        theme(axis.text.x = element_text(angle=25, hjust = 1, size = 8), legend.position = "none",
              axis.title.x = element_blank(),
              plot.margin=unit(c(0.1,0.2,0.05,0.2), "cm")) +
        geom_rect(data=recessions.trim, aes(xmin=Peak, xmax=Trough, ymin=-Inf, ymax=+Inf), fill='grey', alpha=0.2) 

# Plot a histogram with a normal density overlapping
## Source: https://stackoverflow.com/questions/6967664/ggplot2-histogram-with-normal-curve
n        <- nrow(dfData01)
mean     <- mean(dfData01[, "Schmidt"])
sd       <- sd(dfData01[,"Schmidt"])
binwidth <- 0.3 
df       <- data.frame(x = rnorm(n, mean, sd))
dfData03 <- cbind(dfData01, df)

p1 <- ggplot(dfData03, aes(x = Schmidt, mean = mean, sd = sd, binwidth = binwidth, n = n)) +
      geom_histogram(binwidth = binwidth, 
        colour = "#A43E53", fill = "#E9975C", size = 0.1, alpha = 0.4) +
      stat_function(fun = function(x) dnorm(x, mean = mean, sd = sd) * n * binwidth,
          color = "#0E1139", size = 1)+
      labs(title="", y = "Count", 
                  x= "Rocket Portfolio Returns (%)") +
     theme(axis.text.x = element_text(size = 6),   axis.title.x = element_text(size = 6),
           axis.text.y = element_text(size = 6),   axis.title.y = element_text(size = 6),
           plot.margin=unit(c(-1,-0.5,-2,-0.5), "cm"))+
      theme_bw()

# Make a QQplot
p2 <- ggplot(dfData01, aes(sample = Schmidt))+
  stat_qq(colour = "#E9975C", fill = "#E9975C", size = 1.5, alpha = 0.4) +
  stat_qq_line(color = "#0E1139", size = 1)+
  labs(title="", y = "Sample\n Quantile", 
                  x= "Theoretical Quantile")+
  theme(axis.text.x = element_text(size = 6),   axis.title.x = element_text(size = 6),
           axis.text.y = element_text(size = 6),   axis.title.y = element_text(size = 6),
        plot.margin=unit(c(-1,-0.5,-2,-0.5), "cm"))+
  theme_bw()

#png(file = "Series_01.png", width = 450, height = 300)
grid.arrange(p0, grid.arrange(p1, p2, nrow=1), nrow = 2)
#dev.off()

# Cumulative moment plot
dfMoment <- data.frame(matrix(NA, ncol = 5, nrow = nrow(dfData01)))
names(dfMoment) <- c("Date", "First", "Second", "Third", "Fourth")

# This will compute the cumulative moments (centered on zero)
# up to order k
for (k in 1:4) {
  for (i in 1:nrow(dfData01)) {
    dfMoment[i,k+1] <- mean(dfData01$Schmidt[1:i]^k)
  }
}

dfMoment$Date      <- dfData01$Date
dfMomentstd        <- dfMoment

for (j in 2:5){
  dfMomentstd[,j]  <- (dfMoment[,j]-mean(dfMoment[,j]))/sd(dfMoment[,j])
}

dfMomentstd[1, 2]  <- 0

dfData02           <- melt(dfMomentstd, id.vars = "Date")
names(dfData02)    <- c("DATE", "Variable", "Value")

p3 <- ggplot(dfData02) +
        geom_line(aes(x = DATE, y = Value, color = Variable), alpha = 0.9, size = 0.9)+
        labs(y = "Std. cum. moment", x= "", color = "Moment") +
        scale_x_date(date_breaks = "12 months", labels=date_format("%Y"))+ 
        scale_colour_manual(values = c("#E9975C", "#6D304B", "#A43E53", "#0E1139"), labels = c("First", "Second", "Third", "Fourth")) +
        theme_bw() +
        theme(axis.text.x = element_text(angle=25, hjust = 1, size = 8), legend.position = "bottom",
              axis.title.x = element_blank(),
              plot.margin=unit(c(0.1,0.2,0.05,0.2), "cm"))

#png(file = "Series_02.png", width = 450, height = 150)
p3
#dev.off

sd(dfData01$Schmidt)
mean(dfData01$Schmidt)
```


# Question 1 - what is the maximum amount of asset (in mln USD) that can be invested in the risky portfolio?

* Regulator: with holding the invested portfolio for 100 days, the Value-at-Risk (VaR) at 99.9% level multiplied by a multiplier 3 has to be lower than the capital of the bank
* Capital of the bank 30 million

```{r, eval = TRUE, message = FALSE, warning = FALSE}
# Computes the Hill estimator
# Reference: https://github.com/bpfaff/evir/blob/master/R/eda.R
# had to adapt so it will spit the confidence interval
fhill <- 
function(data, option = c("alpha","xi","quantile"), start = 15, end = NA,
    reverse = FALSE, p = NA, ci = 0.95, auto.scale = TRUE, labels = TRUE, plot = FALSE)
{
    data <- as.numeric(data)
    ordered <- rev(sort(data))
    ordered <- ordered[ordered > 0]
    n <- length(ordered)
    option <- match.arg(option)
    if((option == "quantile") && (is.na(p)))
        stop("Input a value for the probability p")
    if((option == "quantile") && (p < 1 - start/n)) {
	cat("Graph may look strange !! \n\n")
	cat(paste("Suggestion 1: Increase `p' above",
                  format(signif(1 - start/n, 5)), "\n"))
	cat(paste("Suggestion 2: Increase `start' above ",
                  ceiling(length(data) * (1 - p)), "\n"))
    }
    
    k <- 1:n
    loggs <- logb(ordered)
    avesumlog <- cumsum(loggs)/(1:n)
    xihat <- c(NA, (avesumlog - loggs)[2:n])
    alphahat <- 1/xihat
    y <- switch(option,
	    alpha = alphahat,
	    xi = xihat,
	    quantile = ordered * ((n * (1 - p))/k)^(-1/alphahat))
    ses <- y/sqrt(k)
    if(is.na(end)) end <- n
    x <- trunc(seq(from = min(end, length(data)), to = start))
    y <- y[x]
    ylabel <- option
    yrange <- range(y)
    if(ci && (option != "quantile")) {
       	qq <- qnorm(1 - (1 - ci)/2)
       	u <- y + ses[x] * qq
       	l <- y - ses[x] * qq
       	ylabel <- paste(ylabel, " (CI, p =", ci, ")", sep = "")
       	yrange <- range(u, l)
    }
    if(option == "quantile") ylabel <- paste("Quantile, p =", p)
    index <- x
    if(reverse) index <-  - x
    if(plot == TRUE){
      if(auto.scale)
          plot(index, y, ylim = yrange, type = "l", xlab = "", ylab = "",
  	     axes = FALSE, ...)
      else plot(index, y, type = "l", xlab = "", ylab = "", axes = FALSE, ...)
      axis(1, at = index, labels = paste(x), tick = FALSE)
      axis(2)
      threshold <- findthresh(data, x)
      axis(3, at = index, labels = paste(format(signif(threshold, 3))),
           tick = FALSE)
      box()
      if(ci && (option != "quantile")) {
         	lines(index, u, lty = 2, col = 2)
         	lines(index, l, lty = 2, col = 2)
      }
      if(labels) {
         	title(xlab = "Order Statistics", ylab = ylabel)
         	mtext("Threshold", side = 3, line = 3)
      }
    }
    invisible(list(x = index, y = y, u=u, l=l))
}

lHill <- fhill(dfData01$Schmidt, option = c("alpha","xi","quantile"), start = 15,
end = NA, reverse = FALSE, p = NA, ci = 0.95,
auto.scale = TRUE, labels = TRUE, plot = FALSE)

dfHill <- data.frame(lHill$x, lHill$y, lHill$l, lHill$u)
names(dfHill) <- c("statistics", "Estimate", "Lower", "Upper")

p4 <- ggplot(dfHill, aes(x = statistics, y = Estimate)) +
        geom_line(aes(x = statistics, y = Estimate), colour = "#6D304B", alpha = 0.9, size = 0.9)+
        geom_line(aes(x = statistics, y = Lower), colour = "#E9975C", alpha = 0.9, size = 0.7, linetype="dashed")+
        geom_line(aes(x = statistics, y = Upper), colour = "#E9975C", alpha = 0.9, size = 0.7, linetype="dashed")+
        geom_ribbon(aes(ymin = Lower, ymax = Upper), fill = "#E9975C", alpha = 0.2) +
        scale_x_continuous(breaks = seq(from = min(dfHill$statistics), to = max(dfHill$statistics), by = 75)) +
        theme_bw() +
        labs(y = "Hill estimator", x= "Order statistics") +
        theme(axis.text.x = element_text(size = 7),   axis.title.x = element_text(size = 8),
           axis.text.y = element_text(size = 7),   axis.title.y = element_text(size = 8),
           plot.margin=unit(c(0.1,0.2,0.05,0.2), "cm"), panel.grid.minor.x = element_blank())

#png(file = "Series_03.png", width = 450, height = 150)
p4
#dev.off

lHill <- fhill(dfData01$Schmidt, option = c("alpha","xi","quantile"), start = 165,
end = 172, reverse = FALSE, p = NA, ci = 0.95,
auto.scale = TRUE, labels = TRUE, plot = FALSE)


dfHill <- data.frame(lHill$x, lHill$y, lHill$l, lHill$u)
names(dfHill) <- c("statistics", "Estimate", "Lower", "Upper")

p5 <- ggplot(dfHill, aes(x = statistics, y = Estimate)) +
        geom_line(aes(x = statistics, y = Estimate), colour = "#6D304B", alpha = 0.9, size = 0.9)+
        geom_line(aes(x = statistics, y = Lower), colour = "#E9975C", alpha = 0.9, size = 0.7, linetype="dashed")+
        geom_line(aes(x = statistics, y = Upper), colour = "#E9975C", alpha = 0.9, size = 0.7, linetype="dashed")+
        geom_ribbon(aes(ymin = Lower, ymax = Upper), fill = "#E9975C", alpha = 0.2) +
        scale_x_continuous(breaks = seq(from = min(dfHill$statistics), to = max(dfHill$statistics), by = 2)) +
        theme_bw() +
        labs(y = "Hill estimator", x= "Order statistics") +
        theme(axis.text.x = element_text(size = 7),   axis.title.x = element_text(size = 8),
           axis.text.y = element_text(size = 7),   axis.title.y = element_text(size = 8),
           plot.margin=unit(c(0.1,0.2,0.05,0.2), "cm"), panel.grid.minor.x = element_blank())


#png(file = "Series_04.png", width = 150, height = 75)
p5
#dev.off

```


```{r, eval = TRUE, message = FALSE, warning = FALSE}
kstar = 170
alphastar <- dfHill[which(dfHill$statistics == kstar), "Estimate"]
print(alphastar)


```


